---
title: "Practicle Machine learning project"
author: "Vinh Nguyen"
date: "Sunday, June 21, 2015"
output: html_document
---
**Introduction**

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it.

In this project, the goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

**Data**

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har.

**loading training and testing dataset**

```{r}
training <- read.csv("pml-training.csv",na.strings=c("NA",""))
testing <-read.csv("pml-testing.csv",na.strings=c("NA",""))
```
First look of training data
```{r}
dim(training)
```
The outcome is “classe” variable
```{r}
table(training$classe)
```
The train dataset has 160 variables 19622 observations.

**remove missing values, is.na() or “”**

```{r}
var <- names(training)[apply(training,2,function(x) table(is.na(x))[1]==19622)]   
train2<- training[,var]
test2 <- testing[,c(var[-length(var)],names(testing)[length(testing)])] # test dataset no classe variable
```
**discards unuseful predictors**

Only considering numeric variable from HAR sensor
```{r}
removeIndex <- grep("timestamp|X|user_name|new_window|num_window",names(train2))
train3 <- train2[,-c(removeIndex, length(train2))]
test3  <- test2[,-c(removeIndex, length(test2))]

```
**Check the near Zero covariates and correlation matrix**

removing zero covariates

```{r}
library(caret)
```
```{r}
nzv <- nearZeroVar(train3, saveMetrics=TRUE)
nzv
```
```{r}
nzv[nzv$nzv,]
```
**Remove highly correlated covariates.**
```{r}
corrM <- cor(train3)
library(corrplot)
corrplot(corrM, method="circle",tl.cex=0.5)
```
```{r}
highCorr <- findCorrelation(corrM, cutoff = .75)     # high correlation
train4<-cbind(classe=train2$classe,train3[,-highCorr])    
test4 <- test3[, -highCorr]        # dataframe of test predictors
```
**Split training dataset into training/testing for model evaulation**

```{r}
set.seed(1234)
inTrain = createDataPartition(train4$classe, p = 3/4)[[1]]
trainPart = train4[ inTrain,]
testPart =  train4[-inTrain,]
```
**Random Forest algorithm to predict.**

```{r}
library(randomForest)
```
```{r}
rfModel <- randomForest(classe ~ .,data = trainPart,importance = TRUE,ntrees = 500)
print(rfModel)
```
```{r}
par(mar=c(3,4,4,4))                               
plot(rfModel)  
```
**Test sample and cross validation**


```{r}
out.test<-predict(rfModel,test4) 
out.test[1:20]
```

**k fold cross validataion, which takes too much computing time PCA with threshold 0.80, two misclassified in the first 20.**

```{r}
preProc <- preProcess(trainPart[,-1], method="pca", thresh=0.8)
trainPC <- predict(preProc, trainPart[,-1])
#modPC <- train(trainPart$classe~., method="rf", data=trainPC)
modPC <- randomForest(trainPart$classe~., data=trainPC, importance=TRUE, ntree=10)

testPC <- predict(preProc, testPart[,-1])
out.testPC<-predict(modPC, newdata=testPC)
table(out.testPC, testPart$classe)
```

```{r}
testPC <- predict(preProc, test4)
out.testPC<-predict(modPC, newdata=testPC)

out.testPC[1:20]  # the 1st and 3rd mis-classified
```
```{r}
out.test[1:20]==out.testPC[1:20]
```
**Saving the output**

```{r}
answers<- as.vector(out.test[1:20])
#answers = rep("A", 20)
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(answers)
```


